---
title: "Data Manipulation with `dplyr`"
output: github_document
---

Once you've imported data, you're going to need to do some cleaning up. 

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collaps = TRUE,
  fig.width = 6,
  fig.asp = 6,
  out.width = "90%"
)
# changes how it knits to markdown
```

## Overview
```{r}
library(tidyverse)

options(tibble.print_min = 3)

litters_data <- read_csv("./data/FAS_litters.csv")
litters_data <- janitor::clean_names(litters_data)

pups_data <- read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data <- janitor::clean_names(pups_data)
```

### `select`

You can specify the columns you want to keep by naming all of them:

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

You can specify a range of columns to keep:

```{r}
select(litters_data, group:gd_of_birth)
```

You can also specify columns you’d like to remove:

```{r}
select(litters_data, -pups_survive)
```

You can rename variables as part of this process:

```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

If all you want to do is rename something, you can use `rename` instead of `select`. This will rename the variables you care about, and keep everything else:

```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)

```

There are some handy helper functions for select; read about all of them using ?select_helpers. I use starts_with(), ends_with(), and contains() often, especially when there variables are named with suffixes or other standard patterns:

```{r}
select(litters_data, starts_with("gd"))
```

I also frequently use is everything(), which is handy for reorganizing columns without discarding anything:

```{r}
select(litters_data, litter_number, pups_survive, everything())
```

`relocate` does a similar thing (and is sort of like rename in that it’s handy but not critical):

```{r}
relocate(litters_data, litter_number, pups_survive)
```


### `filter`
