---
title: "Data Manipulation with `dplyr`"
output: github_document
---

Once you've imported data, you're going to need to do some cleaning up. 

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collaps = TRUE,
  fig.width = 6,
  fig.asp = 6,
  out.width = "90%"
)
# changes how it knits to markdown
```

## Overview
```{r}
library(tidyverse)

options(tibble.print_min = 3)

litters_data <- read_csv("./data/FAS_litters.csv")
litters_data <- janitor::clean_names(litters_data)

pups_data <- read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data <- janitor::clean_names(pups_data)
```

### `select`

You can specify the columns you want to keep by naming all of them:

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

You can specify a range of columns to keep:

```{r}
select(litters_data, group:gd_of_birth)
```

You can also specify columns you’d like to remove:

```{r}
select(litters_data, -pups_survive)
```

You can rename variables as part of this process:

```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

If all you want to do is rename something, you can use `rename` instead of `select`. This will rename the variables you care about, and keep everything else:

```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)

```

There are some handy helper functions for select; read about all of them using ?select_helpers. I use starts_with(), ends_with(), and contains() often, especially when there variables are named with suffixes or other standard patterns:

```{r}
select(litters_data, starts_with("gd"))
```

I also frequently use is everything(), which is handy for reorganizing columns without discarding anything:

```{r}
select(litters_data, litter_number, pups_survive, everything())
```

`relocate` does a similar thing (and is sort of like rename in that it’s handy but not critical):

```{r}
relocate(litters_data, litter_number, pups_survive)
```


### `filter`

Some ways you might filter the litters data are:

* `gd_of_birth == 20`
* `pups_born_alive >= 2`
* `pups_survive != 4`
* `!(pups_survive == 4)`
* `group %in% c("Con7", "Con8")`
* `group == "Con7" & gd_of_birth == 20`

A very common filtering step requires you to omit missing observations. You can do this with filter, but I recommend using `drop_na` from the `tidyr` package:

* `drop_na(litters_data)` will remove any row with a missing value
* `drop_na(litters_data, wt_increase)` will remove rows for which `wt_increase` is missing


### `mutate`

```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
  # wt_gain_kg = wt_gain * 2.2
```


### `arrange`
In comparison to the preceding, arranging is pretty straightforward. You can arrange the rows in your data according to the values in one or more columns:

```{r}
head(arrange(litters_data, group, pups_born_alive), 10)
```


### `%>%`

We can (1) define intermediate datasets (or overwrite data at each stage)
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = 
  mutate(
    litters_data_selected_cols, 
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
litters_data_with_vars_without_missing = 
  drop_na(litters_data_with_vars, wt_gain)
litters_data_with_vars_without_missing
```

OR we can (2) nest function calls
```{r}
litters_data_clean = 
  drop_na(
    mutate(
      select(
        janitor::clean_names(
          read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
          ), 
      -pups_survive
      ),
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
    ),
  wt_gain
  )

litters_data_clean
```

*NOTE*: These are both confusing and bad: the first gets confusing and clutters our workspace, and the second has to be read inside out.

Piping solves this problem. It allows you to turn the nested approach into a sequential chain by passing the result of one function call as an argument to the next function call:
```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)

litters_data
```

Using piping when not using a function in tidyverse
```{r}
litters_data %>%
  lm(wt_gain ~ pups_born_alive, data = .) %>%
  broom::tidy()
```
This doesn't operate on the "tibble in, tibble out" idea, but you can still use it. 
Putting a '.' for data is telling R to take the previous thing and place it there. You can still use non-tidyverse functions with the piping as long as you use the period as the marker. `broom::tidy()` takes the output and makes it prettier... 
