---
title: "Data Import"
output: github_document
---

```{r, echo = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

## Data Import: CSVs

Let's import data using the `readr` package.
```{r}
litters_df <- read_csv("data/FAS_litters.csv")
litters_df <- janitor::clean_names(litters_df)
# converts all names to snake case
```
NOTE: Give relative paths - that's why R projects are so helpful. You can start giving relative paths from the R project and send the R project so the code works for everyone. 

Janitor Package: we don't need this entirely; we just need the `clean_names` function. So, we just call on the package when we need it. Don't have to load it completely. 
`clean_names` converts all variable names to snake case. 


#### First thing to do: *Look at the data*

```{r}
litters_df

head(litters_df)
tail(litters_df)
```

```{r, eval = FALSE}
view(litters_df)
```

```{r}
skimr::skim(litters_df)
```


#### `read_csv` options ...
Setting your missing values:
```{r}
read_csv("data/FAS_litters.csv", na = c("", "NA", 999, 88))
```

Skip the first two rows:
```{r}
read_csv("data/FAS_litters.csv", skip = 2)
```

`col_types`:
Update a column type if the dataset was incorrectly considered 'character' or 'double' (numeric) when imported


## Other file formats

We need to read in an excel spreadsheet ...

```{r}
mlb_df <- read_excel("data/mlb11.xlsx")
```

```{r, eval = FALSE}
view(mlb_df)
```

```{r}
lotr_words_df = 
  read_excel(
    "data/LotR_Words.xlsx",
    range = "B3:D6")
```


## Still more formats

Read in a SAS dataset.

```{r}
pulse_df <- read_sas("data/public_pulse_data.sas7bdat")
```


## Data export

Using relative paths in the same way that you import.
```{r}
write_csv(lotr_words_df, file = "data/lotr_words_df.csv")
```

If you want to add to a new folder > you need to create the new folder first.
```{r}
write_csv(lotr_words_df, file = "results/lotr_words_df.csv")
```


## Why not base r???
```{r}
dont_do_this_df <- read.csv("data/FAS_litters.csv")
```

`read_csv` vs `read.csv`: `read.csv` is not a good thing to use (weird columns, etc.)


## Gitignore

Can use this to specify if there are certain files / datasets that you don't want to include in github. 
Use for data that you don't ever want to save and update on github. 