---
title: "Data Import"
output: github_document
---

```{r, echo = FALSE}
library(tidyverse)
```

## Data Import: CSVs

Let's import data using the `readr` package.
```{r}
litters_df <- read_csv("data/FAS_litters.csv")
litters_df <- janitor::clean_names(litters_df)
# converts all names to snake case
```
NOTE: Give relative paths - that's why R projects are so helpful. You can start giving relative paths from the R project and send the R project so the code works for everyone. 

Janitor Package: we don't need this entirely; we just need the `clean_names` function. So, we just call on the package when we need it. Don't have to load it completely. 
`clean_names` converts all variable names to snake case. 


#### First thing to do: *Look at the data*

```{r}
litters_df

head(litters_df)
tail(litters_df)
```

```{r, eval = FALSE}
view(litters_df)
```

```{r}
skimr::skim(litters_df)
```


#### `read_csv` options ...
Setting your missing values:
```{r}
read_csv("data/FAS_litters.csv", na = c("", "NA", 999, 88))
```

Skip the first two rows:
```{r}
read_csv("data/FAS_litters.csv", skip = 2)
```

`col_types`:
Update a column type if the dataset was incorrectly considered 'character' or 'double' (numeric) when imported


## Other file formats

