---
title: "Tidy Data"
output: github_document
date: "2022-09-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
options(tibble.print_min = 5)
```

### Using `pivot_longer`
```{r}
pulse_df <- haven::read_sas('data/public_pulse_data.sas7bdat') %>%
  janitor::clean_names()
```
Data shows ID, age at baseline, sex, and depression scores. Data is not tidy.
We want to `pivot_longer`, take those columnns, stack them on top of each other, and create a new variable that indicates what visit we're at.

```{r}
pulse_tidy_data <-
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi",
    names_prefix = "bdi_score_"
  )
```
`names_prefix` --> drops whatever you list in quotes.

### Doing things in one swoop + `mutate`
```{r}
pulse_df <- haven::read_sas('data/public_pulse_data.sas7bdat') %>%
  janitor::clean_names() %>% 
    pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi",
    names_prefix = "bdi_score_"
  ) %>% 
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)
  ) %>% 
  arrange(id, visit)
```


### Learning Assessment: Litters Data
```{r}
litters_df <- read_csv('data/FAS_litters.csv') %>% 
  janitor::clean_names()

litters_tidy_df <- litters_df %>% 
  select(litter_number, gd0_weight, gd18_weight) %>% 
  pivot_longer(
    c(gd0_weight, gd18_weight),
    names_to = "gd",
    values_to = "weight"
  ) %>% 
  mutate(
    gd = recode(gd, "gd0_weight" = 0, "gd18_weight" = 18))
```
Answer: Yes, this is relatively tidy. But this makes it harder to compute weight change, so it may not be as useful. 
Need to think about what you want to do. Sometimes this can be better, sometimes you'll want it long. 


### `pivot_wider`
Creating tibble
```{r}
analysis_results <- tibble(
  group = c('treatment', 'treatment', 'placebo', 'placebo'),
  time = c('pre', 'post', 'pre', 'post'),
  mean = c(4, 8, 3.5, 4)
)
```

What if we want to change this into wide format? 
```{r}
analysis_results_wide <- pivot_wider(
  analysis_results,
  names_from = 'time',
  values_from = 'mean'
)
```

`names_from` = Column names are coming from the variable `time`
`values_from` = Values come from the variable listed here that will move into the new variable


### Lord of the Rings Exercise

